---
title: "3_Exercise2"
author: "Haifaa Alzahrani"
date: "11/22/2020"
output: html_document
---

## [Exercise 2](https://misk-data-science.github.io/misk-homl/docs/99x2-portfolio-builder.html)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# Import packages
# Helper packages
library(tidyverse)
library(dplyr)     # for data manipulation
library(ggplot2)   # for awesome graphics
#library(corrplot)  # for correlation plot

# Modeling process packages
library(rsample)   # for resampling procedures
library(caret)     # for resampling and model training
#library(h2o)       # for resampling and model training

# Results
library(Metrics) 
```

```{r}
# Read the dataset
netFlow <- read_csv('Data/IoT_Intrusion_Dataset_2020_Prepared.csv')

# Examine the dataset
#The dimensions
glimpse(netFlow)
dim(netFlow)
```

```{r}
netFlow <- netFlow %>% 
  select(-Flow_ID, -X1)
```


```{r}
# We have to encode the labels into numeric data
netFlow$Label <- factor(netFlow$Label)
netFlow$Label <- as.numeric(netFlow$Label)
```

```{r}
set.seed(123)  # for reproducibility
netFlow <- sample_n(netFlow, 500)
split  <- rsample::initial_split(netFlow, prop = 0.7, 
                                 strata = "Label")
netFlow_train  <- rsample::training(split)
netFlow_test   <- rsample::testing(split)
```

### **Q1** Depending on the type of response variable, apply a linear or logistic regression model.
### - First, apply the model to your data without pre-applying feature engineering processes.
```{r}
# Train model using 10-fold cross-validation
set.seed(123)  # for reproducibility
(lm_model <- train(
  form = Label ~ ., 
  data = netFlow_train, 
  method = "lm",
  trControl = trainControl(method = "cv", number = 10)
))
```
### - Create and a apply a blueprint of feature engineering processes that you think will help your model improve.

### **Ans1** Blueprint to pre-process the data
```{r}
library(recipes)

blueprint <- recipe(Label ~ ., data = netFlow_train) %>%
  step_nzv(all_nominal())  %>%
  step_integer(matches("Src_IP|Dst_IP|Timestamp|Label")) %>%
  #step_center(all_numeric(), -all_outcomes()) %>%
  #step_scale(all_numeric(), -all_outcomes()) %>%
  #step_pca(all_numeric(), -all_outcomes())
  step_dummy(all_nominal(), -all_outcomes(), one_hot = TRUE)

blueprint
```

```{r}
prepare <- prep(blueprint, training = netFlow_train)
```

```{r}
baked_train <- bake(prepare, new_data = netFlow_train)
baked_test <- bake(prepare, new_data = netFlow_test)
```

### - Now reapply the model to your data that has been feature engineered.
### - Did your model performance improve?
### **Ans1**
```{r}
# Train model using 10-fold cross-validation
set.seed(123)  # for reproducibility
(lm_model <- train(
  Label ~ ., 
  data = baked_train, 
  method = "lm",
  trControl = trainControl(method = "cv", number = 10)#,
  #preProcess = c("nzv") #, "center", "scale"
))   #blueprint
```

### **Q2** Apply a principal component regression model.
### - Perform a grid search over several components.
### - Identify and explain the performance of the optimal model.
```{r}
# perform 10-fold cross validation on a PCR model tuning the 
# number of principal components to use as predictors from 1-20
set.seed(123)
cv_model_pcr <- train(
  Label ~ ., 
  data = netFlow_train, 
  method = "pcr",
  trControl = trainControl(method = "cv", number = 10),
  preProcess = c("zv", "center", "scale"),
  tuneLength = 100
  )

# model with lowest RMSE
cv_model_pcr$bestTune
```


```{r}
# results for model with lowest RMSE
cv_model_pcr$results %>%
  dplyr::filter(ncomp == pull(cv_model_pcr$bestTune))
```

### **Q3** Apply a partial least squares regression model.
### - Perform a grid search over several components.
### - Identify and explain the performance of the optimal model.
```{r}
# perform 10-fold cross validation on a PLS model tuning the 
# number of principal components to use as predictors from 1-30
set.seed(123)
cv_model_pls <- train(
  Label ~ ., 
  data = netFlow_train, 
  method = "pls",
  trControl = trainControl(method = "cv", number = 10),
  preProcess = c("zv", "center", "scale"),
  tuneLength = 30
)

# model with lowest RMSE
cv_model_pls$bestTune
```
```{r}
# results for model with lowest RMSE
cv_model_pls$results %>%
  dplyr::filter(ncomp == pull(cv_model_pls$bestTune))
```

### **Q4** Apply a regularized regression model.
### - Perform a grid search across alpha parameter values ranging between 0–1.
### - What is the optimal alpha and lambda values?
### - What is the MSE and RMSE for this optimal model?
### - How does it compare to your previous models?
### **Ans4** It is not applicable for the classification problems. 

