---
title: "3_Exercise3"
author: "Haifaa Alzahrani"
date: "11/22/2020"
output: html_document
---

## [Exercise 3](https://misk-data-science.github.io/misk-homl/docs/99x3-portfolio-builder.html)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# Import packages
# Helper packages
library(tidyverse)
library(dplyr)     # for data manipulation
library(ggplot2)   # for awesome graphics

# Modeling process packages
library(rsample)   # for resampling procedures
library(caret)     # for resampling and model training
library(ranger)   # for fitting Random Forest

# Results
library(Metrics) 
```

```{r}
# Read the dataset
netFlow <- read_csv('Data/IoT_Intrusion_Dataset_2020_Prepared.csv')

# Examine the dataset
#The dimensions
glimpse(netFlow)
dim(netFlow)
```

```{r}
netFlow <- netFlow %>% 
  select(-Flow_ID, -X1)
```

```{r}

netFlow <- rename(netFlow, c("Flow_Byts_s" = "Flow_Byts/s", 
                  "Flow_Pkts_s" = "Flow_Pkts/s",
                  "Bwd_IAT_Mean_1" = "Bwd_IAT_Mean.1",
                  "Fwd_Pkts_s" = "Fwd_Pkts/s",
                  "Bwd_Pkts_s" = "Bwd_Pkts/s",
                  "Down_Up_Ratio" = "Down/Up_Ratio",
                  "Fwd_Byts_b_Avg" = "Fwd_Byts/b_Avg",
                  "Fwd_Pkts_b_Avg" = "Fwd_Pkts/b_Avg", 
                  "Bwd_Byts_b_Avg" = "Bwd_Byts/b_Avg",
                  "Bwd_Pkts_b_Avg" = "Bwd_Pkts/b_Avg")) 
```

```{r}
glimpse(netFlow)
```


```{r}
# We have to encode the labels into numeric data
netFlow$Label <- factor(netFlow$Label)
netFlow$Label <- as.numeric(netFlow$Label)
```

```{r}
set.seed(123)  # for reproducibility
netFlow <- sample_n(netFlow, 500)
split  <- rsample::initial_split(netFlow, prop = 0.7, 
                                 strata = "Label")
netFlow_train  <- rsample::training(split)
netFlow_test   <- rsample::testing(split)
```

### **Q1** Apply a MARS model with all features.
### - How does the model performance compare to your previous models?
### - How many of the features are influential? Which 10 features are considered most influential?
### - Does your model include hinge functions? If so, explain their coefficient and plot their impact on the predicted response variable.
### - Does your model include interactions? If so, pick the interaction effect that is most influential and explain the coefficient.
### **Ans1** It is not applicable for the classification problems. 

### **Q12** Apply a random forest model.
### - First, apply a default random forest model.

```{r}
# number of features
features <- setdiff(names(netFlow_train), "Label")

# perform basic random forest model
fit_default <- ranger(
  formula    = Label ~ ., 
  data       = netFlow_train, 
  num.trees  = length(features) * 10,
  mtry       = floor(length(features) / 3),
  respect.unordered.factors = 'order',
  verbose    = FALSE,
  seed       = 123
  )

#Default results are based on OOB errors:
# look at results
fit_default

# compute RMSE (RMSE = square root of MSE)
sqrt(fit_default$prediction.error)
```
